---
title: "Dynamic Assessment Report created by the spictapp"
author: "Tobias K. Mildenberger"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_depth: 2
        toc_float:
            collapsed: false
            smooth_scroll: false
        theme: united
        highlight: tango
    pdf_document:
        toc: true
        toc_depth: 2
        highlight: tango
params:
  rv: NA
bibliography: "`r file.path(tempdir(), 'spictapp.bib')`"
header-includes:
 \usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage[table]{xcolor}
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r ReaddataLoadLibraries, message=FALSE, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      include = TRUE,
                      collapse = TRUE,
                      comment = "",
                      fig.show = "hold",
                      fig.align="center",
                      fig.width=10, fig.height=7)
library(spict)
library(knitr)
```

<br><br><br>
\vspace*{1.5cm}

<font size="-3">
\small
**Disclaimer**

This report has been created automatically by the spictapp software. The authors
of the app do not accept any responsibility or liability for the correctness and
reliability of the results summarised in the app or this report nor any
conclusions based on them. The interpretation and use of the spictapp software,
its results, and this report is strictly at your own risk.

**Package version numbers**

This report was created with by spictapp (version `0.9.1`) based on the
**`SPiCT`** package (version `r packageVersion("spict")`; @spict).

</font>
\normalsize

<br><br><br>
\vspace*{1.2cm}


# Data
The results in this report are based on the data set labelled
`r paste0("'",strsplit(params$rv$filename,"Data: ")[[1]][2],"'")`.
The columns in the uploaded data set (Table 1) were used in the
analysis (first 10 rows of each column displayed). The complete data set is
included in the appendix of this report.

```{r rawdat}
if(!is.null(params$rv$dat)) kable(
                                head(params$rv$dat, 10), booktabs = TRUE,
                                caption = "Input data.")
```

The input list used for the SPiCT assessment can also be displayed graphically
with the standard data plot in the spict package as shown in Fig. \ref{fig:dataplot}.

```{r dataplot, fig.width=6, fig.height=7, fig.cap = "SPiCT data plot.\\label{fig:dataplot}"}
if(!is.null(params$rv$inp)) plotspict.data(params$rv$inp)
```

<br><br>

# Priors

Following priors were defined:
`r if(!is.null(params$rv$inp)) paste0(names(params$rv$inp$priors)[which(params$rv$inp$priorsuseflags == 1)], collapse = ", ")`.

```{r priorsplot, fig.width=5.5, fig.height=8, fig.cap = "Priors. \\label{fig:priorsplot}"}
plotspict.priors.inp <- function(x, do.plot=NULL, stamp=get.version(), automfrow=TRUE){
    ## if check.inp has not been run yet
    if(!inherits(x, "spictcls")){
        x <- try(check.inp(x, verbose = FALSE, mancheck = FALSE), silent=TRUE)
        if(inherits(x, "try-error")){
            stop("Provided object 'x' needs to be an input list or a fitted spict object!")
        }
    }
    ## if fitted object
    if(inherits(x, "spictcls") && any(names(x) == "opt")){
        inp <- x$inp
        rep <- x
        isRep <- TRUE
    }else{
        inp <- check.inp(x, verbose = FALSE, mancheck = FALSE)
        rep <- NULL
        isRep <- FALSE
    }
    useflags <- inp$priorsuseflags
    inds <- which(useflags == 1)
    ninds <- length(inds)
    if(!is.null(do.plot)) automfrow <- FALSE
    if(automfrow) {
        nopriors <- get.no.active.priors(inp)
        op <- par(mfrow=n2mfrow(nopriors))
        on.exit(par(op))
    }
    counter <- 0
    nused <- sum(useflags)
    if (ninds > 0){
        for (i in 1:ninds){
            j <- inds[i]
            priorvec <- inp$priors[[j]]
            nm <- names(inp$priors)[j]
            isGamma <- FALSE
            nmpl <- sub('log', '', nm)
            nmpl <- sub('gamma','',nmpl)
            if(isRep){
                par <- get.par(nm, rep, exp=FALSE)
            }
            if (length(grep('gamma', nm)) == 1){
                isGamma <- TRUE
                if(nm=="logngamma" && isRep) par <- get.par("logn",rep,exp=FALSE)
            }

            repriors <- c('logB', 'logF', 'logBBmsy', 'logFFmsy')
            if (nm %in% repriors){
                if(isRep) par <- par[priorvec[5], , drop=FALSE]
                nmpl <- paste0(nmpl, fd(priorvec[4]))
                if (nm == 'logB'){
                    nmpl <- add.catchunit(nmpl, inp$catchunit)
                }
            }
            if(isRep){
                nrowPar <- nrow(par)
            }else
                nrowPar <- 1
            for (rr in 1:nrowPar){
                if (nrowPar > 1){
                    nmpl <- paste0(nmpl, rr)
                }
                prvec <- priorvec
                if(is.list(priorvec)) prvec <- priorvec[[rr]]
                if(isRep){
                    mu <- ifelse(is.na(par[rr, 4]), prvec[1], par[rr, 2])
                    sd <- ifelse(is.na(par[rr, 4]), prvec[2], par[rr, 4])
                }else{
                    mu <- prvec[1]
                    sd <- prvec[2]
                }
                if(isRep){
                    if(isGamma && is.na(par[rr, 4])){
                        xmin <- 1e-12
                        xmax <- qgamma(0.99,shape=mu,rate=sd)
                    } else {
                        xmin <- mu - 3*sd
                        xmax <- mu + 3*sd
                    }
                }else{
                    xmin <- mu - 3*sd
                    xmax <- mu + 3*sd
                }
                xpr <- xpo <- seq(xmin, xmax, length=200)
                if(!isGamma) {
                    priorvals <- dnorm(xpr, prvec[1], prvec[2])
                }  else  {
                    priorvals <- dgamma(xpr, prvec[1], prvec[2])
                }

                if(isRep){
                    if (is.na(par[rr, 4])){
                        posteriorvals <- NULL
                    } else {
                        if(isGamma) xpo <- seq(mu - 3*sd, mu + 3*sd, length=200)
                        posteriorvals <- dnorm(xpo, par[rr, 2], par[rr, 4])
                    }
                }else posteriorvals <- NULL
                plot(exp(xpr), priorvals, typ='l', xlab=nmpl, ylab='Density', log='x',
                     lwd=1.5, ylim=c(0, max(priorvals, posteriorvals)*1.3))
                if(isRep){
                    if (is.na(par[rr, 4])){
                        if (!is.na(par[rr, 2])){
                            abline(v=exp(par[rr, 2]), lty=2, col=3, lwd=1.5)
                        }
                        legend('topright', legend=c('Prior', 'Post. Mean'), lty=1:2,
                               col=c(1, 3), lwd=1.5)
                    } else {
                        lines(exp(xpo), posteriorvals, col=3, lwd=1.5)
                        legend('topright', legend=c('Prior', 'Post.'), lty=1,
                               col=c(1, 3), lwd=1.5)
                    }
                }
                box(lwd=1.5)
                if (isRep && rep$opt$convergence != 0){
                    warning.stamp()
                }
                counter <- counter + 1
                if(!is.null(do.plot) && counter >= do.plot) {
                    txt.stamp(stamp)
                    return()
                }
            }
        }
        txt.stamp(stamp)
    }
}

if(!is.null(params$rv$inp)){
    plotspict.priors.inp(params$rv$inp)
}
```

<br><br>

# Assessment results

The seed value for reproducible results was set to `r params$rv$seed`.

## Convergence

```{r}
if(!is.null(params$rv$fit)){
    if(!params$rv$fit$opt$convergence)
        writeLines(paste0("The SPiCT model reached convergence.")) else
        writeLines(paste0("Subsequent results have to be treated with caution as the model did not reach convergence."))
}else{
    writeLines(paste0("The following sections will be incomplete as no SPiCT model was fitted. Make sure to run 'Fit SPiCT' in the spictapp."))
}
```

## Model parameters

The following table includes the estimates (`estimate`) together with lower and
upper 95% confidence intervals (`cilow` and `ciupp`, respectively), as well as
the log estimates (`log.est`) of all fixed parameters of the SPiCT model.

```{r}
if(!is.null(params$rv$fit))
    kable(
        round(sumspict.parest(params$rv$fit),3),
        booktabs = TRUE,
        caption = "Model parameters.")
```

<font size="-3">
\small Note, that the parameter 'alpha' is only included if a survey index was
included in the input list and might be missing if SPiCT is used with catch and
effort only.

</font>
\normalsize

## Reference points

SPiCT estimates deterministic and stochastic biological reference points, namely
$MSY$, $B_{MSY}$, and $F_{MSY}$. In most cases, the difference between the
deterministic and stochastic reference levels are minor, but the stochastic
reference levels correct for the presence of the noise term in the biomass
equation of SPiCT.

### Stochastic reference points

Estimated stochastic reference points with lower and upper 95% confidence
intervals, the log estimates, and the relative difference to the deterministic
reference points (`rel.diff.Drp`) are:

```{r}
if(!is.null(params$rv$fit))
    kable(
        round(sumspict.srefpoints(params$rv$fit),3),
        booktabs = TRUE,
        caption = "Stochastic reference points.")
```

<font size="-3">
\small Note, that the stochastic reference points are only defined for $n > 1$.
Therefore, SPiCT overwrites the stochastic reference points with the
deterministic reference points.

</font>
\normalsize

### Deterministic reference points

Comparitively, the deterministic reference points are:

```{r}
if(!is.null(params$rv$fit))
    kable(
        round(sumspict.drefpoints(params$rv$fit),3),
        booktabs = TRUE,
        caption = "Deterministic reference points.")
```

## States

The biomass and fishing mortality in absolute terms and relative to reference
points at the time of the last observation (end of catch or effort interval or
time of last index observation) are called 'States' in SPiCT and estimated with
confidence intervals as:

```{r}
if(!is.null(params$rv$fit))
    kable(
        round(sumspict.states(params$rv$fit),3),
        booktabs = TRUE,
        caption = "Model states.")
```

<font size="-3">
\small Note, that the reference points can correspond to the deterministic or
stochastic reference points. This can be chosen within spictapp. If $n < 1$, the
deterministic reference points are used.

</font>
\normalsize

## Predictions

Similarly to the states, the predictions refer to the absolute and relative
reference points at the specified time (chosen with the setting called
'Management evaluation time' or `maneval`). Additionally, this table includes
predicted catch during the management period (as specified with 'Management
interval' or `maninterval`) and `E(B_inf)` is the predicted equilibrium biomass
and current fishing mortality.

```{r fitpred}
if(!is.null(params$rv$fit))
    kable(
        round(sumspict.predictions(params$rv$fit),3),
        booktabs = TRUE,
        caption = "Model predictions.")
```

The 4 main plots of SPiCT summarising the results graphically, are the relative
biomass and fishing mortality (relative to reference points), the catch plot,
and the Kobe plot. The kobe plot is a phase plot showing the stock development
in terms of biomass and fishing mortality simultaneously. The shaded areas show
the 95% confidence intervals.

```{r fitplot, fig.width=9, fig.height=8, fig.cap = "SPiCT plot.\\label{fig:fitplot}"}
if(!is.null(params$rv$fit)) plot2(params$rv$fit)
```

<font size="-3">
\small Find the absolute biomass and fishing mortality graphs among others in
the appendix of this report.

</font>
\normalsize

<br><br>

# Diagnostics

The evaluation and cross-validation of the model fitting is a crucial step in
any stock assessment. The minimum model evaluation for a SPiCT assessment should
include the consideration of the assumptions of the input data (see
@spict_guidelines), checking residuals against violation of independence,
normality, biases, or auto-correlation, as well as evaluating the prior and
posterior distributions of used priors. Additionally, the retrospective analysis
and sensitivity analysis for intital values is recommended.

```{r}
if(!is.null(params$rv$fit))
    kable(sumspict.diagnostics(params$rv$fit),
        booktabs = TRUE,
        caption = "Model diagnostics.")
```

This can also be evaluated visually:

```{r diagplot, fig.width=7, fig.height=8, fig.cap = "Model diagnostics.\\label{fig:diagplot}"}
if(!is.null(params$rv$fit)) plotspict.diagnostic(params$rv$fit)
```

Ideally, the prior and posterior distributions do not fully overlap. Overlapping
distributions indicate that the prior highly affects the model fit, which could
potentially be due to low contrast in the data.

```{r priorplot, fig.width=5.5, fig.height=8, fig.cap = "Priors.\\label{fig:priorplot}"}
if(!is.null(params$rv$fit)) plotspict.priors(params$rv$fit)
```

## Retrospective analysis

The retrospective analysis evaluates the consistency in assessment outcomes as
additional data (data windows) are added or removed. If the estimates of the
retrospective analysis 'do not agree and the disagreement is systematic pattern
of biases rather than a random scattering' the assessment shows the
'retrospective problem' [@mohn1999].

```{r retroplot, fig.width=9, fig.height=8, fig.cap = "Retrospective analysis\\label{fig:retroplot}"}
if(!is.null(params$rv$retro)) plotspict.retro(params$rv$retro)
```

<font size="-3">
\small The differences between the various trajectories should not be too large,
should be within the 95% confidence interals of the base assessment (grey area),
and should not show a consistent pattern, e.g. F consistently smaller the more
data points removed.

</font>
\normalsize

## *Sensitivity analysis to inital values*

The sensitivity analysis evaluates the consistency in assessment outcomes with
different initial values. Disagreeing assessment outcomes for the different
model fits indicate potential problems with underlying assessment.

```{r}
if(!is.null(params$rv$sensi))
    kable(sumspict.ini(params$rv$sensi[[1]]),
        booktabs = TRUE,
        caption = "Relative difference between initial values of different trials and base run.")
```

```{r}
if(!is.null(params$rv$sensi))
    kable(sumspict.ini(params$rv$sensi[[2]]),
        booktabs = TRUE,
        caption = "Initial values (log scale) for the different trials used in the sensitivity analysis.")
```

```{r}
if(!is.null(params$rv$sensi))
    kable(sumspict.ini(params$rv$sensi[[3]]),
        booktabs = TRUE,
        caption = "Paramater estimates for all trials of the sensitivity analysis, as well as the distance from the parameters of each trial to the base run.")
```

<font size="-3">
\small A range of factors can be the reason for high sensitivity to intial
values, such as local minima in the negative loglikelihood function or missing
contrast in the input data.

</font>
\normalsize

<br><br>

# Management

The management scenarios included in spictapp allow comparing and evaluating the
implications of various strategies on predicted catches and stock status in
future years. Following management scenarios were choosen:
`r if(!is.null(params$rv$mana)) names(params$rv$mana$man) else print("No management scenarios analysed!")`

The SPiCT timeline visualises the period of observations, as well as the intermediate
and management period$^*$.

```{r mantimeline}
if(!is.null(params$rv$mana)){
    man.timeline(params$rv$mana)
}else if(!is.null(params$rv$fit)){
    man.timeline(params$rv$fit)
}else{
    if(!is.null(params$rv$inp)) man.timeline(params$rv$inp)
}
```

<font size="-3">
\small $^*$ Note, that the intermediate period defines the time period between
the last observation (end of catch or effort interval or index) and the start of
the management interval and thus, is only present if management does not start
right after the last observation.

</font>
\normalsize

The predicted trajectories of catches and states under various management
scenarios are displayed in different colours in Fig. \ref{fig:manplot}.

```{r manplot, fig.width=9, fig.height=8, fig.cap = "Management scenarios\\label{fig:manplot}"}
if(!is.null(params$rv$mana)) plot2(params$rv$mana)
```

In detial, the difference between the management scenarios are:

```{r mansum}
if(!is.null(params$rv$mana))
    kable(sumspict.manage(params$rv$mana),
        booktabs = TRUE,
        caption = "Management scenarios.")
```

The total allowable catch for each management scenario for the period
`r if(!is.null(params$rv$mana)) params$rv$mana$man[[1]]$inp$maninterval`
is:

```{r mantac}
if(!is.null(params$rv$mana)) lapply(man.tac(params$rv$mana),round,3)
```

<br><br>

# Summary

```{r summary}
if(is.null(params$rv$fit)){
    datapoints <- 0
    datastart <- NaN
    dataend <- NaN
    stockstatusB <- NaN
    bbmsy <- NaN
    stockstatusF <- NaN
    ffmsy <- NaN
    ffac <- "Continuing the F process,"
    manstart <- NaN
    manend <- NaN
    maneval <- NaN
    bpbmsy <- NaN
    predcatch <- NaN
}else{
    datapoints <- length(params$rv$inp$obssrt)
    datastart <- params$rv$inp$timerange[1]
    dataend <- params$rv$inp$timerange[2]
    bbmsy <- round(get.par("logBlBmsy",params$rv$fit, exp = TRUE)[,2],2)
    stockstatusB <- ifelse(bbmsy < 1,"overfished","not overfished")
    ffmsy <- round(get.par("logFlFmsynotS",params$rv$fit, exp = TRUE)[,2],2)
    stockstatusF <- ifelse(ffmsy > 1,"overfished","not overfished")
    manstart <- params$rv$inp$maninterval[1]
    if(round(params$rv$fit$inp$ffac,2) == 1.0){
        ffac <- "Continuing the F process,"
    }else if(params$rv$fit$inp$ffac > 1){
        ffac <- paste0("Increasing the F process by ",round(params$rv$fit$inp$ffac*100), "% in ",manstart,",")
    }else{
        ffac <- paste0("Decreasing the F process by ",round(params$rv$fit$inp$ffac*100), "% in ",manstart,",")
    }
    predcatch <- paste0(round(get.par("logCp",params$rv$fit, exp = TRUE)[,2],2),params$rv$fit$inp$catchunit)
    manend <- params$rv$inp$maninterval[2]
    maneval <- params$rv$inp$maneval
    bpbmsy <- round(get.par("logBpBmsy",params$rv$fit, exp = TRUE)[,2],2)
}

```

The stock assessment with SPiCT is based on `r datapoints` data points
representing the time period from `r datastart` to `r dataend`. The assessment
indicates that the stock is `r stockstatusB` in terms of biomass
($\frac{B}{B_{MSY}} =$ `r bbmsy`), and `r stockstatusF` in terms of fishing
mortality ($\frac{F}{F_{MSY}} =$ `r ffmsy`). `r ffac` the predicted catch for
the period from `r manstart` to `r manend` is `r predcatch` and the predicted
biomass relative to $B_{MSY}$ in `r maneval` is `r bpbmsy`.

<br><br>

# Author's comment

If you have comments or questions please write an [email](mailto:t.k.mildenberger@gmail.com) or post an issue at [GitHub](https://github.com/tokami/apps/issues). You can follow the development of **`SPiCT`** on [ResearchGate](https://www.researchgate.net/project/Stochastic-production-model-in-continuous-time-SPiCT).

<br><br>

# References

<br><br>

# Appendix

## Raw data {#app:data}

The complete data set used for the analysis:

```{r}
if(!is.null(params$rv$dat))
    kable(params$rv$dat,
        booktabs = TRUE,
        caption = "Full uploaded data.")
```

## Full SPiCT summary {#app:summary}

The complete SPiCT model summary:

```{r}
if(!is.null(params$rv$fit)) summary(params$rv$fit)
```

## SPiCT default plots {#app:plots}

```{r fitplotFull, fig.width=9, fig.height=10, fig.cap = "Full SPiCT default plot.\\label{fig:fitplotFull}"}
if(!is.null(params$rv$fit)) plot(params$rv$fit)
```

```{r manplotFull, fig.width=9, fig.height=10, fig.cap = "Management scenarios default SPiCT plot.\\label{fig:maplotFull}"}
if(!is.null(params$rv$mana)) plot(params$rv$mana)
```
