---
title: "Tutorial for spictapp: The Shiny app for the Stochastic Production model in Continuous Time (SPiCT)"
author: "Tobias K. Mildenberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: false
    toc: false
    keep_tex: false
bibliography: bibliography.bib
header-includes: \usepackage{color}
---

```{r settings, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  ", fig.align = 'center', cache=FALSE)
library(png)
library(grid)
```
\vspace*{2cm}

This toturial guides you through a SPiCT assessment using the click-based Shiny
app "spictapp".

```{r home, fig.width=6, echo=FALSE, fig.cap="Home screen of spictapp."}
img <- readPNG("figs/home.png")
 grid.raster(img)
```

\vspace*{1cm}

# Download

The spictapp is hosted on [\underline{GitHub}](https://github.com/tokami/apps/)
and can be downloaded as a [\underline{zip
archive}](https://github.com/tokami/apps/archive/master.zip). Unpack the archive
to the destination of your choice.

\vspace*{1cm}

# Start the App

Before the start of the app, **spictapp** checks if all required R packages are
required and installs any missing packages. To assure windows compatibility
without requiring Rtools (large software package), the binary version of the
spict R package (version 1.3.0) is included in the zip archive and installed
upon start of the app.

The app can be started by double-clicking the respective exectuable in the
spictapp directory, i.e. 'spictapp' for linux and mac and 'spictapp_win' for
windows operating systems (the file endings are '.sh' and '.bat', respectively).

Alternatively the script `runapp.R` in the spictapp directory can be executed
from within R or with `Rscript runapp.R` from the terminal or command line.

\vspace*{1cm}

# Home

On start, the app shows the home screen of spictapp (Fig. 1), which shows four
important plots of a simulated spict assessment (find more information to the
plots below). At the top of the screen is the orange navigation bar of the app,
which guides the user through the indivdual steps of a spict assessment (Fig.
2). The active tab is highlighted in darker tone (tab called 'Load data' in Fig.
2). The tab 'Quit' closes the app and browser window (in any browser other than
firefox >= 46.0.1). This tutorial is structured following the steps of a common
spict assessment and thus the tabs in the navigation bar.

```{r navbar1 , fig.width=8, echo=FALSE, fig.cap="Tabs with assessment steps in the navigation bar."}
img <- readPNG("figs/navbar1.png")
grid.raster(img)
```

At the far right of the navigation bar, the name of the uploaded data set is
displayed (Fig. 3). In this example, the albcaore example data set was selected.
The data name is shown independent on the active tab and helps avoid confusion
when dealing with different data sets.

```{r navbar2, fig.width=0.9, echo=FALSE, fig.cap="Data label in the navigation bar."}
img <- readPNG("figs/navbar2.png")
grid.raster(img)
```

\vspace*{1cm}

# Load data

Any assessment requires input data. SPiCT requires information about the catches
from the commercial fleet with corresponding time intervals and an abundance
index or effort data, alternatively. While the catches and effort refer to an
interval and the times of the start of these intervals have to be provided, the
abdundance indices (multiple indices possible) correspond to a specific point
in time, which can be specified as 2015.37 for mid May for example. Find more
information about the data requirements of **SPiCT** in the two vignettes
[\underline{SPiCT Guidelines}](../apps/www/spict_guidelines.pdf)
and [\underline{SPiCT Handbook}](../apps/www/spict_handbook.pdf).

The button 'Browse...' in the 'Load Data' tab allows you to browse through your
directories and upload any data set to **spictapp** (Fig. 4). The only
requirement for the data file is that it has either the '.txt' or '.csv' file
extension. Both file types can be created from Excel or R. The specific
properties of the file can be changed within the app with the options given for
seperators, quotes, and header. After uploading, the apps displays the data in
its raw format and will try to automatically match the column names with the
names expected by **SPiCT**. If successful in matching expected column names,
the data set will also be displayed under 'Data with assigned columns'.

```{r loadData1, fig.width=6, echo=FALSE, fig.cap="Upload data to spictapp."}
img <- readPNG("figs/loadData1.png")
grid.raster(img)
```

If not all columns expected by **SPiCT** could be matched, the user can select
the columns corresponding to the commercial catch observations ('obsC'), with
corresponding times ('timeC') and index ('obsI') or effort ('obsE') observations
with corresponding times ('timeI') or ('timeE'), respectively, where the names
in brackets reflect the correspinding  standard **SPiCT** variable names (Fig.
5). Note, that the app allows to input several columns for the index
observations and times, but only one column for the catch and effort
observations and times.

```{r loadData2, fig.height=8, echo=FALSE, fig.cap="Assign columns."}
img <- readPNG("figs/loadData2.png")
grid.raster(img)
```

With pressing 'Update data', the columns are assigned to
the corresponding **SPiCT** variables and the resulting data is displyed under
'Data with assigned columns' (Fig. 6).

```{r loadData3, fig.width=6, echo=FALSE, fig.cap="Uploaded data with assigned columns."}
img <- readPNG("figs/loadData3.png")
grid.raster(img)
```

**SPiCT** also allows you to specify the uncertainty around input data as a
factor to multiply estimated observation noise with. For example, the
uncertainty of the catch observations might have changed over time due to an
improved data monitoring system. The three input fields below the 'Update data'
button, let you assign corresponding columns in your data to these variables
(called 'stdevfacC', 'stdevfacI', and 'stdevfacE' in **SPiCT**, respectively).

One of the three originalexample data sets included in the spict pacakge can be
choosen by pressing 'Use example data set?' at the bottom of the page (Fig. 5).
Note that many more example data sets are included in the data directory of the
spictapp zip archive.

\vspace*{1cm}

# Explore & modify data

This step of the analysis visualises the input data and allows to explore and
modify the input data to the SPiCT assessment. The SPiCT timeline visualises the
time period with observations (catch, indices, and/or effort), the forecast
period which is used in all management related functions, and the potential
intermediate period between observations and start of the management (Fig. 7).
Furthermore, the observations are displayed as time series graphs (Fig. 7) and
as text output (Fig. 8). Note that catches in SPiCT are generally labelled with
a capital 'C', indices with a capital 'I', and effort with a capital 'E'. The
number of plots can vary with the number of index time series (here 2).

```{r exploData1, fig.width=6, echo=FALSE, fig.cap="Explore and modify SPiCT input data, part I."}
img <- readPNG("figs/explo1.png")
grid.raster(img)
```

Under 'General Settings', a range of options allows to modify the input data and
model settings (Figures 7 and 8). The Euler discretisation time step defining the
number of time steps within a year (default is 16) can be changed. The timing of
the indices can be adjusted by entering a number(s) for each index, e.g. 0.25 to
adjust the timing of the index to April. The time period of the observations can
be shortened by changing the sliders. Be aware that this affects all time series
(e.g. catches and all indices). For seasonal catch/effort observations, the
number of seasonal, the season type and and spline order can be adjusted.

The 'Management settings' allow to specify the management interval (period for
which to predict catches and in which to apply potential management strategies),
as well as the management evaluation time (time at which to evaluate predicted
states), and the management strategy by changing the fishing mortality, as a
factor to multiply current F by or as an absolute F value. Note, that the
management related settings can be re-adjusted in the 'Management scenarios'
section.

```{r exploData2, fig.width=6, echo=FALSE, fig.cap="Explore and modify SPiCT input data, part II."}
img <- readPNG("figs/explo2.png")
grid.raster(img)
```

SPiCT defines 3 uninformative priors by default: (i) on the shape of the
production curve (`log(n)`); (ii) on the relation of the biomass process error
to the index observation error (`log(alpha)`); and on the fishing mortality
process error to the catch observation error (`log(beta)`). These priors can
easily be switched off or modified (Fig. 9). In addition, SPiCT allows to
specify prior distributions on a range of other model parameters and quantities,
such as the carrying capacity (`log(K)`) or the stock productivity (`log(m)`).
However, caution has to be applied when specifying additional priors as they
affect model estimates and uncertainties and can greatly affect estimated stock
status.

```{r exploData3, fig.width=6, echo=FALSE, fig.cap="Explore and modify SPiCT input data, part III."}
img <- readPNG("figs/explo3.png")
grid.raster(img)
```

Instead of the standard SPiCT data plot, a more advanced version can be
displayed with the option 'Plot the advanced data plot?' (at the bottom of the
sidepanel on the lefthand side; not shown in the screenshots).

\vspace*{1cm}

# Fit SPiCT

After data has been uplaoded and was modified according to meet the assumptions
of the model, SPiCT can be fitted to the input data by clicking on the 'Fit
SPiCT' button (Fig. 10). A notification in the lower right corner of the browser
informs you the model is being fitted, which can take some time dependent on the
number of data points in the input data. After completion, a message will
indicate if the model converged and the 4 main SPiCT graphs are shown (Fig. 10).
Note, that the model results should not be used if the model did not reach
convergence. There are a number potential reasons why the model did not reach
convergence and steps that can be taken subsequently (see the [\underline{SPiCT
Handbook}](../apps/www/spict_handbook.pdf)).

```{r fit1, fig.width=6, echo=FALSE, fig.cap="Fit SPiCT."}
img <- readPNG("figs/fit1.png")
grid.raster(img)
```

Additionally, the summary of the SPiCT fit is shown and the graphs with the
absolute biomass and fishing mortality states as well as the production curve
and either the seasonal F plot or a plot with time to recovery at the bottom of
the page. The seed value allows to generate reproducible results. The optimiser
settings do not need to be changed in the most cases.

\vspace*{1cm}

# Diagnostics

Evaluating the diagnostics of a model fit is as important as the fit itself. The
'Diagnostics' tab allows to compare the posterior and prior distributions (if
priors were specified; Fig. 12).

```{r diag1, fig.width=6, echo=FALSE, fig.cap="Priors."}
img <- readPNG("figs/priors.png")
grid.raster(img)
```

Furthermore, it shows a range of model diagnostics, such as normality of
residuals (Fig. 13).

```{r diag2, fig.width=6, echo=FALSE, fig.cap="Diagnostics."}
img <- readPNG("figs/diag.png")
grid.raster(img)
```

And allows to perform a retrospective and senstivity analysis to
intial values (Fig. 14).

```{r diag3, fig.width=6, echo=FALSE, fig.cap="Run retrospective and sensitivity analysis to intial values."}
img <- readPNG("figs/runretrosensi.png")
grid.raster(img)
```

The results of the retrospective and sensitivity analysis to intial values are
shown under model diagnostics after completion (plot not shown here).

\vspace*{1cm}

# Management scenarios

The implications of different management strategies on the predicted catch and
states can be explored by applying different management scenarios. Any number of
the 8 default management scenarios can be chosen:

\begin{enumerate}
   \item \textbf{currentCatch}: Keep the catch of the current year (i.e. the last observed catch).
   \item \textbf{currentF}:  Keep the F of the current year.
   \item \textbf{Fmsy}: Fish at Fmsy i.e. F=Fmsy.
   \item \textbf{noF}: No fishing, reduce to 1\% of current F.
   \item \textbf{reduceF25}: Reduce F by 25\%.
   \item \textbf{increaseF25}: Increase F by 25\%,
   \item \textbf{msyHockeyStick}: Use ICES MSY hockey-stick advice rule [@msycat34].
   \item \textbf{ices}: Use ICES MSY 35th hockey-stick advice rule [@wklifeix].
\end{enumerate}

By default the scenarios 'currentCatch' and 'Fmsy' are chosen (Fig. 15). The
management interval and management evaluation time can be adjusted. If an
intermediate period is defined, the catch during the intermediate period can be
defined. If undefined, the fishing mortality process is continued in the
intermediate period. By default the median of the predicted catch distribution
is used as the the total allowable catch (TAC), but any value can be chosen and
the TAC will be re-calculated (Fig. 17).

```{r man1, fig.width=6, echo=FALSE, fig.cap="Run management scenarios."}
img <- readPNG("figs/runmanage.png")
grid.raster(img)
```

The management plot shows the implications of the management scenarios on the
relative biomass, fishing mortality, catches and kobe plot (Fig. 16). Note, that
the lines corresponding to the different management strategies are only
depicted in the kobe plot if the catch/effort data is not subannual.

```{r man2, fig.width=6, echo=FALSE, fig.cap="Plot of management scenarios."}
img <- readPNG("figs/manplot.png")
grid.raster(img)
```

Below the management plot the summary of the management scenarios as well as the
TAC of each scenario is shown (Fig. 17).

```{r man3, fig.width=6, echo=FALSE, fig.cap="Results of management scenarios."}
img <- readPNG("figs/mansum.png")
grid.raster(img)
```

\vspace*{1cm}

# Summary

The 'Summary' tab summarises the results of all tabs and allows to download the
results in form of a report (html file), a zip archive with the main results as
tables, a zip archive with all figures, or the complete data used and produced
in spictapp as a 'RData' file (Fig. 18). The RData file includes a list called
'rv' (for reactive values) that includes among others, the input data
(`rv$inp`), the fitted object (`rv$fit`), as well as the retrospective and
sensitivity analysis (`rv$retro` and `rv$sensi`, respectively) and the
management results (`rv$mana`) if applied. It can be loaded in a R session by
`load("spictapp_alldata_X_Y.RData")` where 'X' stands for the name of the data
set and 'Y' for the date when it was downloaded. This data set allows to
reproduce all steps of the spictapp in a plane R session using the spict
package. The description of all individual elements of that list (`names(rv)`)
would exceed the scope of this vignette.

```{r sum, fig.width=6, echo=FALSE, fig.cap="Download options."}
img <- readPNG("figs/summary.png")
grid.raster(img)
```

\vspace*{1cm}

# Other
The 'References' tab lists peer-reviewed articles about SPiCT and other
documentation such as vignettes of the SPiCT package and this tutorial. The
'About' tab lists the version number of the app, instructions on how to download
and start the app, information on how to report bugs or issues, ask questions or
follow the development of the packages. 'Quit' closes the connection to the app
and the browser window (except on firefox >= 46.0.1).

# References
