---
title: "Tutorial for spictapp: The Shiny app for the Stochastic Production model in Continuous Time (SPiCT)"
author: "Tobias K. Mildenberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: false
    toc: false
    keep_tex: false
bibliography: bibliography.bib
header-includes: \usepackage{color}
---

```{r settings, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  ", fig.align = 'center', cache=FALSE)
library(png)
library(grid)
```
\vspace*{2cm}

This toturial guides you through a SPiCT assessment using the click-based Shiny
app "spictapp".

```{r home, fig.width=6, echo=FALSE, fig.cap="Home screen of spictapp."}
img <- readPNG("figs/home.png")
 grid.raster(img)
```

\vspace*{1cm}

# Download

The spictapp is hosted on [\underline{GitHub}](https://github.com/tokami/apps/)
and can be downloaded as a [\underline{zip
archive}](https://github.com/tokami/apps/archive/master.zip). Unpack the archive
to the destination of your choice.

\vspace*{1cm}

# Start the App

Before the start of the app, **spictapp** checks if all required R packages are
required and installs any missing packages. To assure windows compatibility
without requiring Rtools (large software package), the binary version of the
spict R package (version 1.3.0) is included in the zip archive and installed
upon start of the app.

The app can be started by double-clicking the respective exectuable in the
spictapp directory, i.e. 'spictapp' for linux and mac and 'spictapp_win' for
windows operating systems (the file endings are '.sh' and '.bat', respectively).

Alternatively the script `runapp.R` in the spictapp directory can be executed
from within R or with `Rscript runapp.R` from the terminal or command line.

\vspace*{1cm}

# Home

On start, the app shows the home screen of spictapp (Fig. 1), which shows four
important plots of a simulated spict assessment (find more information to the
plots below). At the top of the screen is the orange navigation bar of the app,
which guides the user through the indivdual steps of a spict assessment (Fig.
2). The active tab is highlighted in darker tone (tab called 'Load data' in Fig.
2). The tab 'Quit' closes the app and browser window (in any browser other than
firefox >= 46.0.1). This tutorial is structured following the steps of a common
spict assessment and thus the tabs in the navigation bar.

```{r navbar1 , fig.width=8, echo=FALSE, fig.cap="Tabs with assessment steps in the navigation bar."}
img <- readPNG("figs/navbar1.png")
grid.raster(img)
```

At the far right of the navigation bar, the name of the uploaded data set is
displayed (Fig. 3). In this example, the albcaore example data set was selected.
The data name is shown independent on the active tab and helps avoid confusion
when dealing with different data sets.

```{r navbar2, fig.width=0.9, echo=FALSE, fig.cap="Data label in the navigation bar."}
img <- readPNG("figs/navbar2.png")
grid.raster(img)
```

\vspace*{1cm}

# Load data

Any assessment requires input data. SPiCT requires information about the catches
from the commercial fleet with corresponding time intervals and an abundance
index or effort data, alternatively. While the catches and effort refer to an
interval and the times of the start of these intervals have to be provided, the
abdundance indices (multiple indices possible) correspond to a specific point
in time, which can be specified as 2015.37 for mid May for example. Find more
information about the data requirements of **SPiCT** in the two vignettes
[\underline{SPiCT Guidelines}](../apps/www/spict_guidelines.pdf)
and [\underline{SPiCT Handbook}](../apps/www/spict_handbook.pdf).

The button 'Browse...' in the 'Load Data' tab allows you to browse through your
directories and upload any data set to **spictapp** (Fig. 4). The only
requirement for the data file is that it has either the '.txt' or '.csv' file
extension. Both file types can be created from Excel or R. The specific
properties of the file can be changed within the app with the options given for
seperators, quotes, and header. After uploading, the apps displays the data in
its raw format and will try to automatically match the column names with the
names expected by **SPiCT**. If successful in matching expected column names,
the data set will also be displayed under 'Data with assigned columns'.

```{r loadData1, fig.width=6, echo=FALSE, fig.cap="Upload data to spictapp."}
img <- readPNG("figs/loadData1.png")
grid.raster(img)
```

If not all columns expected by **SPiCT** could be matched, the user can select
the columns corresponding to the commercial catch observations ('obsC'), with
corresponding times ('timeC') and index ('obsI') or effort ('obsE') observations
with corresponding times ('timeI') or ('timeE'), respectively, where the names
in brackets reflect the correspinding  standard **SPiCT** variable names (Fig.
5). Note, that the app allows to input several columns for the index
observations and times, but only one column for the catch and effort
observations and times.

```{r loadData2, fig.height=8, echo=FALSE, fig.cap="Assign columns."}
img <- readPNG("figs/loadData2.png")
grid.raster(img)
```

With pressing 'Update data', the columns are assigned to
the corresponding **SPiCT** variables and the resulting data is displyed under
'Data with assigned columns' (Fig. 6).

```{r loadData3, fig.width=6, echo=FALSE, fig.cap="Assign columns."}
img <- readPNG("figs/loadData3.png")
grid.raster(img)
```

**SPiCT** also allows you to specify the uncertainty around input data as a
factor to multiply estimated observation noise with. For example, the
uncertainty of the catch observations might have changed over time due to an
improved data monitoring system. The three input fields below the 'Update data'
button, let you assign corresponding columns in your data to these variables
(called 'stdevfacC', 'stdevfacI', and 'stdevfacE' in **SPiCT**, respectively).

One of the three originalexample data sets included in the spict pacakge can be
choosen by pressing 'Use example data set?' at the bottom of the page (Fig. 5).
Note that many more example data sets are included in the data directory of the
spictapp zip archive.

\vspace*{1cm}

# Explore & modify data

Blabla sPiCT timeline and data plot

```{r exploData1, fig.width=6, echo=FALSE, fig.cap="Assign columns."}
img <- readPNG("figs/loadData3.png")
grid.raster(img)
```
changing dteuler, timing of index catch unit

change time range of all data, robust estimation for catches (several for
indices possible / required if several indices)

seaosnal spict if seaosnal catches, number of seasons, by default dependent on
input data 4 for quarterly catches and 1 for annual catches, season type and
splieorder for season type 1 and 3

mangement interval, management evaluation time, ffac or fcon

default priors

additional priors

display options (advance data plot)

\vspace*{1cm}

# SPiCT assessment

fit spict by click on button,

convergence message

set seed value for reproducible results and assessments

optimiser (expert see handbook)

\vspace*{1cm}

# Diagnostics

run retrospecitve analysis

run sesnitivity analysis to intital values

\vspace*{1cm}

# Management with SPiCT

run different management scenarios, intermediate period, catch during
intermediate period, fratile of catch dist

\vspace*{1cm}

# Summary

all results summarised and options to download (assessment report), all tables
as zip archive, all figures as zip archive or all data as RData (includes a R
list object called 'rv') which includes all used data and estimations allowing
to reproduce the whole functionality of the app in R. The description of all
individual elements of that list would exceed the scope of this vignette. see
the names of all elements with `names(rv)`.

\vspace*{1cm}

# Other

## References

references relevant for this app.

## About

more information about the app, version, developers. question/issues links to
user forum on RG

# References
