---
title: "Tutorial for spictapp: The click-based app for SPiCT"
author: "Tobias K. Mildenberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::html_vignette:
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: false
    toc: false
    keep_tex: false
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{spictapp tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  ", fig.align = 'center', cache=FALSE)
```

This toturial guides you through a SPiCT assessment using the click-based Shiny
app "spictapp".


# Installation

The spictapp requires the SPiCT R package to be installed on your computer. This
package can be downnloaded and installed from GitHub using the `devtools`
package:

```{r, eval=FALSE}
if (!require("devtools"))
  install.packages("devtools")
devtools::install_github("DTUAqua/spict/spict")
```

The spictapp is also hosted on GitHub and can thus be downaloaded and installed by:

```{r, eval=FALSE}
if (!require("devtools"))
  install.packages("devtools")
devtools::install_github("tokami/apps/spictapp")
```

Alternatively, the package can also be installed by the specific exectuable for
your operating system, available [here](https://github.com/tokami/apps).


# Start the App


# Input data


# Explore data


# SPiCT assessment


# Diagnostics


# Management with SPiCT


# References
