---
title: "Dynamic Assessment Report created by ShinyTropFish"
author: "Tobias K. Mildenberger"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_depth: 2
        toc_float:
            collapsed: false
            smooth_scroll: false
        theme: united
        highlight: tango
    pdf_document:
        toc: true
        toc_depth: 2
        highlight: tango
    word_document:
        toc: true
        toc_depth: 2
params:
  shinyRes: NA
header-includes:
 \usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage[table]{xcolor}
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r ReaddataLoadLibraries, message=FALSE, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = FALSE,
                      warning = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      include = TRUE,
                      collapse = TRUE,
                      comment = "",
                      fig.show = "hold",
                      fig.align="center",
                      fig.width=10, fig.height=7)

library(TropFishR)
library(knitr)
library(flextable)
library(pander)

doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

<br><br><br>
\vspace*{1.5cm}

<font size="-3">
\small
**Disclaimer**

This report has been created automatically by the ShinyTropFish software. The
authors of the app do not accept any responsibility or liability for the
correctness and reliability of the results summarised in the app or this report
nor any conclusions based on them. The interpretation and use of the
ShinyTropFish software, its results, and this report is strictly at your own
risk.

**Package version numbers**

This report was created with by ShinyTropFish (version `0.9.1`) based on the
**`TropFishR`** package (version `r packageVersion("TropFishR")`; Mildenberger
et al., 2017).

</font>
\normalsize

<br><br><br>
\vspace*{1.2cm}

# Data

```{r dataset}
dataset <- paste0("'",strsplit(params$shinyRes$filename,"Data: ")[[1]][2],"'")
if(is.null(params$shinyRes$lfqre)){
    datapoints <- 0
    datastart <- NA
    dataend <- NA
    sampletimes <- 0
    datayears <- "-"
    reflevyears <- NA
    MA <- 0
}else{
    datapoints <- as.integer(sum(colSums(as.matrix(params$shinyRes$lfqre$catch))))
    datastart <- format(min(params$shinyRes$lfqre$dates),"%b %d, %Y")
    dataend <- format(max(params$shinyRes$lfqre$dates),"%b %d, %Y")
    sampletimes <- ncol(params$shinyRes$lfqre$catch)
    datayears <- params$shinyRes$years
    reflevyears <- params$shinyRes$yearsRefLev
    MA <- params$shinyRes$ma
}
```

The results in this report are based on the data set labelled `r dataset`. The
columns from the uploaded data set shown in Table 1 were used in the analysis
(first 10 rows displayed). The complete data set is included in the appendix of
this report.

<font size="-3">
\small

```{r rawdat}
if(!is.null(params$shinyRes$dat)){
    if(doc.type == "docx"){
        flextable(
            head(params$shinyRes$dat, 10))
    }else{
        kable(
            head(params$shinyRes$dat, 10), booktabs = TRUE,
            caption = "First 10 rows of the input data.")
    }
}
```

</font>
\normalsize

The length frequency (LFQ) data spans following year(s): `r datayears`. It
comprises `r datapoints` length measurements, which are aggregated over
`r sampletimes` sampling times (Figure 1). The figure shows the raw LFQ data
(Fig. 1a) and after restructuring with among others a moving average of `r MA`
(Fig. 1b).

```{r data_plot, out.width = "80%", fig.width = 9, fig.height = 8, fig.cap = "Raw (a) and restructured (b) length frequency data"}
if(!is.null(params$shinyRes$lfq)){
    par(mar=c(5,5,2,1), mfrow= c(2,1))
    plot(params$shinyRes$lfq, Fname = "catch")
    mtext("a)",adj=0, font=2, line=.2)
    plot(params$shinyRes$lfqre, Fname = "rcounts")
    mtext("b)",adj=0, font=2, line=.2)
}
```


<br><br>
\vspace*{1cm}


# Growth

The Electronic LEngth Frequency ANalysis (ELEFAN; Pauly and David, 1981) allows
the estimation of the parameters of the (seasonalised) von Bertalanffy growth
equation: the asymptotic length $L_{\infty}$ (Linf), the growth coefficient $K$,
and the anchor point $t_a$ (ta). The quantity $\phi_L$ (phiL) is the growth
performance index ($\phi_L = log_{10}(K) + 2 log_{10}(L_{\infty})$). The
quantity Rn is the score value of the ELEFAN fitting procedure. It can be used
to compare fits for the same data set with different settings, but not between
different data sets or the same data set with different bin sizes. The data used
for ELEFAN spans the time period from `r datastart ` to `r dataend`. Estimated
values for the data set `r dataset` are given in Table 2.

<br>
\vspace*{5mm}


```{r growth_parameters}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$resGA)){
    tmp <- as.data.frame(c(lapply(params$shinyRes$parsGrowth,round,3),
                           list(Rn_max = round(params$shinyRes$resGA@fitnessValue,3))))
    names(tmp) <- replace(names(tmp), names(tmp)=="Rn_max", "Rn")
    names(tmp) <- replace(names(tmp), names(tmp)=="phiL", "phiL")
    if(doc.type == "docx"){
        flextable(tmp)
    }else{
        kable(tmp,
              booktabs = TRUE,
              caption = "Estimated growth parameters.")
    }
}
```

<br>
\vspace*{5mm}

Figure 2 shows the score function of the ELEFAN_GA fitting procedure over time
(generations) and thus, allows to evaluate the sufficiency of the number of
generations in finding optimum parameters.

```{r elefan_score_plot, out.width = "75%", fig.width = 8, fig.height = 6, fig.cap = "ELEFAN score plot"}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$resGA)){
    par(mar=c(5,5,2,1))
    plot(params$shinyRes$resGA)
}
```

<br>
\vspace*{8mm}

Figure 3 depicts estimated growth curves fit through the restructured LFQ data.

```{r growth_plot, out.width = "80%", fig.width = 8, fig.height = 5.5, fig.cap = "Length frequency data with fitted growth curves"}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$lfqFC)){
par(mar=c(5,5,2,1))
plot(params$shinyRes$lfqFC, Fname = "rcounts")
lfqFC <- params$shinyRes$lfqFC
tmp <- lfqFitCurves(lfq = lfqFC,
             par=as.list(params$shinyRes$parsGrowth),
             draw = TRUE)
}
```

<br>
\vspace*{8mm}

Estimated growth curves allow to allocate the difference length classes at
different times to potential cohorts (Fig. 4).

```{r cohort_plot, out.width = "80%", fig.width = 8, fig.height = 5.5, fig.cap ="Length frequency data with interpolated cohorts in different colours."}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$lfq)){
    par(mar=c(5,5,2,1))
    lfqc <- lfqCohort(params$shinyRes$lfqFC, calc_dt = FALSE)
    plot(lfqc, Fname = "catch",
         ylim = c(0, max(params$shinyRes$lfq$midLengths)), image.col=NA)
    pal <- colorRampPalette(c(4,5,7,2))
    with(lfqc, image(x = dates, y = midLengths, z = t(cohort),
                     col = adjustcolor(pal(max(cohort, na.rm = TRUE)), 0.75),
                     add=TRUE
                     ))
}
```

Given the growth parameters the LFQ data can be extrapolated backwards onto the
time axis and such indicate the relative recruitment pattern (Table 3 and Fig.
5). The pattern can not be interpreted in absolute terms as information about
the length at age 0 is lacking ($L_0$).

<br>
\vspace*{5mm}


```{r recruitment_table}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$lfq)){
    lfq <- params$shinyRes$lfq
    lfq$par <- params$shinyRes$parsGrowth
    parsRecruit <- recruitment(lfq, plot=FALSE)
    tmp <- data.frame("relative month" = as.integer(parsRecruit$mids),
                      counts = parsRecruit$counts,
                      density = round(parsRecruit$density,3))
    colnames(tmp) <- c("Relative month", "Number", "Density")
    if(doc.type == "docx"){
        flextable(tmp)
    }else{
        kable(tmp,
              booktabs = TRUE,
              caption = "Relative recruitment pattern.")
    }
}
```

<br>
\vspace*{5mm}

```{r recruitment_plot, out.width = "80%", fig.width = 8, fig.height = 5, fig.cap ="Relative recruitment pattern."}
if(!is.null(params$shinyRes$parsGrowth) & !is.null(params$shinyRes$lfq)){
    lfq <- params$shinyRes$lfq
    lfq$par <- params$shinyRes$parsGrowth
    par(mar=c(5,5,2,1))
    tmp <- recruitment(lfq, plot=TRUE)
}
```

<br><br>
\vspace*{1cm}


# Mortality & Selectivity

The length-converted catch curve (LCCC; Pauly, 1990) allows the estimation of
the mortality rates and gear selectivity parameters. Estimated quantities for
the year(s) `r reflevyears` are given in Table 4. GOTCHA describes an alternative
catch curve approach, which aggregates the data by cohorts rather than length
classes. If applied, the results are included in Table 4.

<br><br>
\vspace*{5mm}

```{r mort_pars}
if(!is.null(params$shinyRes$parsMort)){
        tmp <- as.data.frame(round(t(params$shinyRes$parsMort),3))
        if(!is.null(params$shinyRes$parsMortGOTCHA)){
            tmp <- as.data.frame(rbind(tmp, c(round(t(params$shinyRes$parsMortGOTCHA),3),NaN,NaN)))
            rownames(tmp) <- c("LCCC","GOTCHA")
        }
    if(doc.type == "docx"){
        flextable(tmp)
    }else{
        kable(tmp,
              booktabs = TRUE,
              caption = "Mortality and selectivity parameters based on the length-converted catch curve.")
    }
}
```

<br><br>
\vspace*{5mm}

The catch curve with the selected data points for the regression analysis is
shown in Figure 6.

```{r catch_curve, out.width = "90%", fig.width = 8, fig.height = 5, fig.cap ="Length-converted catch curve."}
if(!is.null(params$shinyRes$parsMort)){
    par(mar=c(5,5,2,1))
    plotLCCC(params$shinyRes$resLCCC)
}
```

<br><br>
\vspace*{8mm}

Based on the non-selected data points in the catch curve, the trawl-like gear
selectivity can be estimated (Fig. 7).

```{r catch_curve_sel, out.width = "90%", fig.width = 8, fig.height = 5, fig.cap="Selectivity curve based on the length-converted catch curve."}
if(!is.null(params$shinyRes$parsMort)){
    par(mar=c(5,5,2,1))
    plotLCCC_sel(params$shinyRes$resLCCC)
}
```

<br><br>
\vspace*{8mm}

```{r catch_curve_GOTCHA, out.width = "90%", fig.width = 8, fig.height = 5, fig.cap ="GOTCHA catch curve."}
if(!is.null(params$shinyRes$parsMortGOTCHA)){
    par(mar=c(5,5,2,1))
    plotLCCC(params$shinyRes$resGOTCHA)
}
```

<br><br>
\vspace*{1cm}

# Reference levels

The estimation of the biological reference levels is based on the length-based
yield per recruit model (Thompson and Bell, 1934). Estimated levels for the
year(s) `r reflevyears` are given in Table 5.

<br><br>
\vspace*{5mm}

```{r ypr_pars}
if(!is.null(params$shinyRes$parsRef)){
    tmp <- as.data.frame(round(params$shinyRes$parsRef,3))
    if(doc.type == "docx"){
        flextable(tmp)
    }else{
        kable(tmp,
              booktabs = TRUE,
              caption = "Referenve levels.")
    }
}
```

<br><br>
\vspace*{5mm}

```{r figNum_calc}
if(is.null(params$shinyRes$parsMortGOTCHA)){
    figNumYPR1 <- 8
    figNumYPR2 <- 9
    figNumBPR <- 10
}else{
    figNumYPR1 <- 9
    figNumYPR2 <- 10
    figNumBPR <- 11
}
```

Figure `r figNumYPR1` shows the yield per recruit curve, where the black and
blue solid lines show the yield and biomass per recruit for a range of fishing mortality
values, respectively. The colourful dashed lines show the reference levels, where
green depicts the $F_{0.1}$ reference level. The black dashed line indicates the
current fishing mortality based on the catch curve analysis.

```{r ypr_plot1, out.width = "85%", fig.width = 9, fig.height = 6, fig.cap="Yield per recruit curve."}
if(!is.null(params$shinyRes$parsRef)){
    par(mar=c(5,5,2,4))
    plotYPR(params$shinyRes$resYPR, mark=TRUE)
}
```

<br><br>
\vspace*{8mm}

Figure `r figNumYPR2` shows the yield per recruit isopleth graph. The colour
indicates the yield per recruit from low (blue) to high (red) values. The black
solid lines are the isopleths indicating different areas of the same yield. The
black dashed line indicates the current fishing mortality and selectivity based
on the catch curve analysis.

```{r ypr_plot2, out.width = "65%", fig.width = 9, fig.height = 7, fig.cap="Yield per recruit isopleth graph."}
if(!is.null(params$shinyRes$parsRef)){
    par(mar=c(5,5,2,1))
    plotYPR(params$shinyRes$resYPR_Lc, mark=TRUE,xaxis1 = "FM", yaxis_iso = "L50")
}
```

<br><br>
\vspace*{8mm}

Figure `r figNumBPR` shows the biomass per recruit isopleth graph. The colour
indicates the biomass per recruit from low (red) to high (blue) values. The
black solid lines are the isopleths indicating different areas of the same
biomass. The black dashed line indicates the current fishing mortality and
selectivity based on the catch curve analysis.

```{r bpr_plot, out.width = "65%", fig.width = 9, fig.height = 7, fig.cap="Biomass per recruit isopleth graph."}
if(!is.null(params$shinyRes$parsRef)){
    par(mar=c(5,5,2,1))
    plotYPR(params$shinyRes$resYPR_Lc, mark=TRUE, yaxis1 = "B_R",xaxis1 = "FM", yaxis_iso = "L50")
}
```

<br><br>
\vspace*{1cm}


# Summary

```{r summary}
if(is.null(params$shinyRes$dat) || is.null(params$shinyRes$parsMort) || is.null(params$shinyRes$parsRef)){
    exploi <- NaN
    stockstatusExploi <- NaN
    fishmort <- NaN
    f01 <- NaN
    ff01 <- NaN
    ff01Ind <- NaN
    stockstatusFF01 <- NaN
}else{
    exploi <- round(params$shinyRes$parsMort[which(names(params$shinyRes$parsMort) == "F")] /
                    params$shinyRes$parsMort[which(names(params$shinyRes$parsMort) == "Z")], 2)
    stockstatusExploi <- ifelse(exploi >= 0.5,"overfished","not overfished")
    fishmort <- round(params$shinyRes$parsMort[which(names(params$shinyRes$parsMort) == "F")], 2)
    f01 <- round(params$shinyRes$parsRef[which(names(params$shinyRes$parsRef) == "F01")], 2)
    ff01 <- round(params$shinyRes$parsMort[which(names(params$shinyRes$parsMort) == "F")]/
                  params$shinyRes$parsRef[which(names(params$shinyRes$parsRef) == "F01")] ,2)
    ff01Ind <- ifelse(ff01 >= 1, "larger", "smaller")
    stockstatusFF01 <- ifelse(ff01 > 1,"overfished","not overfished")
}
```

The estimation of growth parameters using ELEFAN with the genetic algorithm as
included in the R package **`TropFishR`** (Mildenberger et al., 2017) is based on
`r datapoints` length measurements representing the time period from `r datastart`
to `r dataend`. Based on estimated growth parameters, two dependent
stock status for the year(s) `r reflevyears` indicators were estimated by means
of the length-converted catch curve and the length-based yield per recruit
model. The exploitation rate (`r paste0("F/Z = ", exploi)`) indicates that the
stock is `r stockstatusExploi` based on the length-converted catch curve. The
current fishing mortality (`r paste0("F = ", fishmort)`) based on the catch
curve is `r ff01Ind` than the reference level (`r pander(paste0("$F_{0.1}$ = ", f01))`)
based on the yield per recruit analysis and thus indicates that the stock is `r stockstatusFF01`
(`r pander(paste0("$F/F_{0.1}$ = ", ff01))`).

<br><br>

# Author's comment

If you have comments or questions please write an
[email](mailto:t.k.mildenberger@gmail.com) or post an issue at
[GitHub](https://github.com/tokami/TropFishR/issues). You can follow the
development of **`TropFishR`** on
[ResearchGate](https://www.researchgate.net/project/TropFishR).


<br><br>

# References

Mildenberger, T. K., Taylor, M. H., & Wolff, M. (2017). TropFishR: an R package
for fisheries analysis with length-frequency data. Methods in Ecology and
Evolution, 8(11), 1520-1527.

Pauly, D. (1990). Length-converted catch curves and the seasonal growth of
fishes. Fishbyte, 8(3), 33-38.

Pauly, D., & David, N. (1981). ELEFAN I, a BASIC program for the objective
extraction of growth parameters from length-frequency data. Meeresforschung,
28(4), 205-211.

Thompson, W. F., & Bell, F. H. (1934). Effect of changes in intensity upon total
yield and yield per unit of gear. Report of the International Fisheries
Commission, 8, 7-49.

<br><br>


# Appendix

## Raw data {#app:data}

The complete data set used for the analysis is given below.

<font size="-4">
\small

```{r full_data}
if(!is.null(params$shinyRes$dat)){
    tmp <- as.data.frame(params$shinyRes$dat)
    if(doc.type == "docx"){
        flextable(tmp)
    }else{
        kable(tmp,
              booktabs = TRUE,
              caption = "Full uploaded data.")
    }
}
```
</font>
\normalsize
